Combined annotation exploration for SCPCP000006

### 2025-03-17 analyses/cell-type-wilms-tumor-06/notebook/07_combined_annotation_across_samples_exploration.Rmd
Apply suggestions from code review
Co-authored-by: Joshua Shapiro <josh.shapiro@ccdatalab.org>

From the pre-processed and labeled `Seurat` object from `results/{sample_id}/06_infercnv_HMM-i3_{sample_id}_reference-{reference}.rds`, we will extract the following information per cell:

- the predicted compartment in `fetal_kidney_predicted.compartment` and `fetal_kidney_predicted.compartment.score`
- the predicted cell type in `fetal_kidney_predicted.cell_type` and `fetal_kidney_predicted.cell_type.score`
- the sample identifier in `sample_id`
- the CNV `{dupli|loss}` estimation per chromosome {i} and arm `{p|q}`: `has_{dupli|loss}_chr{i}{p|q}`, in 1 to 22
- the `counts` of [markers genes](https://github.com/AlexsLemonade/OpenScPCA-analysis/tree/main/analyses/cell-type-wilms-tumor-06#the-table-celltype_metadatacsv-contains-the-following-column-and-information)

@@

We only allow a bit of flexibility in terms of CNV profile for immune and endothelium cells that have a high predicted score. 
Indeed, we know that false positive CNV can be observed in a cell type specific manner. 
Please note that in `06_infercnv.R`, we renamed all `immune` and `endothelium` cells with a `fetal_kidney_predicted.compartment.score` \> `r params$predicted.celltype.threshold` as `normal`.

@@

- *cancer stroma*: We define as *cancer stroma* all cancer cells from the stroma compartment.
- *blastema*: We defined as *blastema* every cancer cell that has a `fetal_kidney_predicted.cell_type == mesenchymal cell`. 
We know that these *mesenchymal* cells are cells from the cap mesenchyme that are not expected to be in a mature kidney. 
These blastema cells should express higher `CITED1` and/or `NCAM1`.
- *cancer epithelium*: We defined as *cancer epithelium* all cancer cells that are neither stroma nor blastemal cells. 
We expect these cells to express epithelial markers. 
Their predicted cell type should correspond to more mature kidney epithelial subunits.

@@

*VIM* do not seem to be neither specific nor universal.

@@

*THY1* do not seem to be neither specific nor universal.

@@

(\*) To the best of our knowledge, there is no Wilms tumor specific and universal (i.e. expressed by all cells in every sample) marker of blastema cells. 
We expect however blastema cells to express higher levels of stemness markers such as `CITED1`, `SIX1/2`, `NCAM1`.

@@

Blastema cells are enriched in `CITED1+` cells, but some samples do not express it at all.

@@

`NCAM1` seems to be more cancer specific and broadly expressed among samples.

@@

As expected, epithelial cells express higher expression of `CDH1`.

@@

- The mean CNV profile of the Wilms tumor cohort seems reasonable in regards to the one published in [Cresswell et al, bioRxiv (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1). 
We identify some commonalities but also some divergence. 
They can be due to technical limitation of our CNV inference but can also reflect some specificities of this cohort (e.g. pre/post treatment samples). 
It would be really useful to compare for some sample the `infercnv` result to a ground truth such as SNP Array (to be requested).

>>>>

From the pre-processed and labeled `Seurat` object in `results/{sample_id}/06_infercnv_HMM-i3_{sample_id}_reference-{reference}.rds`, we will extract the following per-cell information:

- the predicted compartment in `fetal_kidney_predicted.compartment` and `fetal_kidney_predicted.compartment.score`
- the predicted cell type in `fetal_kidney_predicted.cell_type` and `fetal_kidney_predicted.cell_type.score`
- the sample identifier in `sample_id`
- the CNV `{dupli|loss}` estimate per chromosome `{i}`, where `i` is in `{1-22}`, and arm `{p|q}`: `has_{dupli|loss}_chr{i}{p|q}`
- the `counts` of [markers genes](https://github.com/AlexsLemonade/OpenScPCA-analysis/tree/main/analyses/cell-type-wilms-tumor-06#the-table-celltype_metadatacsv-contains-the-following-column-and-information)

@@

We only allow a bit of flexibility in terms of CNV profile for immune and endothelium cells that have a high predicted score. 
Indeed, we know that false positive CNV can be observed in a cell type specific manner. 
Please note that in `06_infercnv.R`, we renamed all `immune` and `endothelium` cells with a `fetal_kidney_predicted.compartment.score` > `r params$predicted.celltype.threshold` as `normal`.

@@

- *cancer stroma*: We define as *cancer stroma* all cancer cells from the stroma compartment.
- *blastema*: We defined as *blastema* every cancer cell that has a `fetal_kidney_predicted.cell_type == mesenchymal cell`. 
We know that these *mesenchymal* cells are cells from the cap mesenchyme that are not expected to be in a mature kidney. 
These blastema cells should express higher *CITED1* and/or *NCAM1*.
- *cancer epithelium*: We defined as *cancer epithelium* all cancer cells that are neither stroma nor blastemal cells. 
We expect these cells to express epithelial markers. 
Their predicted cell type should correspond to more mature kidney epithelial subunits.

@@

*VIM* expression does not seem to be either specific or universal.

@@

*THY1* expression does not seem to be either specific or universal.

@@

(\*) To the best of our knowledge, there is no Wilms tumor specific and universal (i.e. expressed by all cells in every sample) marker of blastema cells. 
We expect however blastema cells to express higher levels of stemness markers such as *CITED1*, *SIX1/2*, *NCAM1*.

@@

Blastema cells are enriched in *CITED1+* cells, but some samples do not express it at all.

@@

*NCAM1* seems to be more cancer specific and broadly expressed among samples.

@@

As expected, epithelial cells express higher expression of *CDH1*.

@@

- The mean CNV profile of the Wilms tumor cohort seems reasonable in regards to the one published in [Cresswell et al. (2024)](https://www.biorxiv.org/content/10.1101/2024.09.04.610994v1). 
We identify some commonalities but also some divergence. 
They can be due to technical limitation of our CNV inference but can also reflect some specificities of this cohort (e.g. pre/post treatment samples). 
It would be really useful to compare for some sample the `infercnv` result to a ground truth such as SNP Array (to be requested).
